The genomic impacts of drift and selection for hybrid performance in maize

### 2015-06-30 response.tex

\documentclass[onecolumn,oneside,letterpaper]{article}

\usepackage{response}
%\usepackage{geneticsT2}

\usepackage{times}
\usepackage{color}
\usepackage{hyperref}

% rayout %
\addtolength{\oddsidemargin}{-2.75cm}
\addtolength{\evensidemargin}{-0.75cm}

\addtolength{\textwidth}{5.5cm}
\addtolength{\topmargin}{-3cm}
\addtolength{\textheight}{4.5cm}

%\parindent=4em
%\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10.0pt}
%\renewcommand{\itemindent}{50pt}
\setlength{\leftskip}{0.5cm}
\renewcommand{\labelitemi}{$-$}

\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
\newcommand{\jristop}{\textcolor{red}{\hline \bf JRI STOPPED HERE\\ \hline}}
\newcommand{\st}[1]{\textcolor{red}{ #1}}
\newcommand{\comst}[1]{\textcolor{red}{ \em{\scriptsize  (#1)}} }
\definecolor{mattgreen} {rgb} {0,0.6,0}
\newcommand{\mbh}[1]{\textcolor{mattgreen}{ \em{\scriptsize  #1}} }
\definecolor{peterpurple}{rgb}{.6,0,.6}
\newcommand{\plr}[1]{\textcolor{peterpurple}{ \emph{\scriptsize (#1)}} }

\renewcommand{\textfraction}{0.01}
\renewcommand{\topfraction}{0.99}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.90}
\renewcommand{\dbltopfraction}{0.95}
\renewcommand{\dblfloatpagefraction}{0.80}
\renewcommand{\sfdefault}{phv}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[CE,CO]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\fancypagestyle{plain}{
	\fancyhf{}
}

\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{,}

% space of double hline in Table
\doublerulesep = 0.4pt

\title{Response to Reviews}

\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{,}

\usepackage{amsmath}

\usepackage{graphicx}

\begin{document}

\maketitle

Stephen's comments:

The reviewers make a number of important points that should be considered in revision. Of particular note, Reviewer 1 highlights the importance of considering the possible effects of purging during the experiment, and this intersects with reviewer 2's comments about lack of details of the experiment and what was being selected, and what the outcome of the experiment was. Could purging, for example, have occurred over the course of the experiment, and did yield improve within lines in addition to increased yield in hybrids? Reviewer 2 also notes a lack of details on the simulation approach; I agree with this, and also thought this was perhaps confusing wording on page 14: "Selections were executed at random".

On the statistical side, candidate selected regions are picked out and effectively 1-sided tests are reported for the rare allele that has increased in frequency. While the authors are reasonably cautious about these cases anyway, at least some attempt to report an experiment-wide p value or FDR for these regions seems useful.

Do the results provide sufficient power to reject the overdominance model? How strong would the signal be likely to be?

One additional comment: the authors note that heterozygosity increases early in the experiment, and discuss various possible explanations for this. But couldn't directional selection on rare alleles be the driver of this, i.e. early on there will be a rise in allele frequencies, increasing diversity, and then subsequently a loss due to fixations?

Reviews:
Reviewer 1 Reviewer's Comments to Author...
I really enjoyed reading this manuscript.  I think this is a well done study 
that provides novel insight in to basic processes in breeding of maize.  Many of 
these results have implications for other species as well.  I have a number of 
minor comments below that I think should be addressed but these do not diminish 
my enthusiasm for the manuscript.

The abstract could use some editing.  There are a number of references in this 
abstract and the abstract is not well focused on the specific findings of this 
study.

In the introduction the authors describe the concept of heterotic groups and 
state that these lines maximize performance and heterosis when crossed.  I would 
caution that this should be clarified to heterosis for yield rather than for all 
traits. Research by other groups (ie Flint-Garcia et al 2009) suggests that 
heterosis can vary for different traits and that the definition of heterotic 
groups is primarily based on yield heterosis, not heterosis for other traits.

The introduction references several studies that looked at genome-wide 
population structure in breeding programs from the Ross-Ibarra group.  I 
hesitate to even make this suggestion (do to potential flaws) but it might be 
appropriate to also include Jiao et al nat Genetics 2012 here as this study also 
looked at a large dataset of SNPs in modern maize.

I was quite intrigued by the difference in distance between cycles 4 and then 8-
16.  The authors suggest that this could be due to some differences in genetic 
maintenance of the populations.  I am curious if there were any particular 
environmental influences during that time period that may have led to stronger 
selection or if there were any differences in agronomic practices that have 
occurred during the long-term experiment (ie irrigation, fertilization, plant 
density, etc).  it would be helpful to know if any of these factors may have 
contributed.

It was a bit surprising to me that there was no differentiation among the 
founder inbreds of the two populations.  If PCA was run only on the founders 
(without any other lines) do you get separation of the two groups or not?

Figure 3 is quite useful and helps to show large physical regions that have lost 
heterozygosity.  The authors discuss the physical-genetic difference.  It would 
be useful to have a supplemental figure that shows these same values (cycle 16 
heterozygosity on y-axis) versus genetic position on x-axis.  I am curious if 
there are ever regions of multiple cM on chromosome arms that become fixed.  
This would be a nice complementary figure.  These are provided for some 
chromosomes in 4B and 5B but would be useful to see genome wide.

The large regions of fixation are quite intriguing.  Is there any evidence for 
major flowering time or disease resistance QTL in these regions?

I am somewhat baffled as to why the Fig S1 was put in supplemental materials.  A 
central aspect of this study is this issue of genetic drift relative to 
selection and there are two pages of results text for this section.  However, 
this useful figure is only in supplemental materials.  I would strongly 
encourage moving this figure to the actual manuscript.  If space is an issue 
then only use figu 4 or fig 5 as these are quite redundant with each other in 
terms of conceptual views. 
The legend to fig S1 could use additional detail.  I am guessing that the black 
lines are 95% CI but this is not clear.  Please explain the black lines.

Figure S2 is not ever referenced in the text as far as I could find.  The data 
table of the same results is referenced but I don’t think the figure is.

The last paragraph of the results section assesses the fixation of particular 
haplotypes over time.  I think it could be very useful to visualize this data 
for each of the large blocks.  This could be done by using a graph of allele 
frequency over time for each region.  Right now these results are presented in 
text form for only two regions but it would be very interesting to see these 
results for all eight regions shown in table 1.  What is the fixation at each 
cycle in the observed data and how often was this region fixed in simulations.  
I appreciate the authors point that these fixations do occur at low rates in 
simulations and since they are studying a genome-wide pattern a number of low 
frequency events are expected to occur by chance.

The authors interpret the rate of heterozygosity in terms of selection and 
drift.  I am assuming that they are interpreting selection to mean the selection 
for advantageous alleles in hybrid test combinations.  However, I wonder about 
the inadvertent selection to maintain heterozygosity?  In work from McMullen and 
Buckler it was argued that the gain in vigour due to heterozygosity of 
pericentromeric regions was larger than expected and that this may be due to 
increased ability to cross and maintain these plants.  Was this concept included 
in the modeling of the data?  In other words, while drift will tend to increase 
homozygosity in a population there may be counteracting forces of heterosis 
effects that will maintain heterozygosity in the same population.  Just curious 
about this concept.

The authors spend some time in the discussion regarding models of heterosis and 
extrapolation of these results to discuss selection.  While this manuscript 
describes materials that are quite relevant for maize breeding programs I think 
it is important to note that these types of materials are not universal.  In 
this case the synthetic populations are derived from inbred lines.  This is 
important in terms of the nature of deleterious alleles that remain.  Strongly 
deleterious alleles have likely been purged from the founders as each founder 
was an inbred line that had been selected for some level of vigor.  This means 
that the remaining single-locus deleterious alleles will all be of quite minor 
effect and therefore selection is likely to act on alleles of minor effect 
rather than major effect deleterious alleles.  I think that is would be useful 
to remind readers of this attribute of the population.  My guess is that if this 
experiment was done with a maize (or teosinte) population that had never been 
passed through an inbred “bottleneck” that there was be some alleles of major 
effect that would be purged quickly and that the role of drift and selection 
might be slightly different.

Reviewer 2 Reviewer's Comments to Author...
Girke et al. in "The genomic impacts of drift and selection for hybrid performance in maize" report a 
retrospective examination of genomic patterns of diversity in a maize experimental population. The data 
set consists of genotyping of ~38,000 SNPs genotyped in individuals from various phases of the selection 
experiment. While the manuscript consists primarily of population genetic analysis, the authors attempted 
to better integrate a popular maize genetic map with the current reference genome.

The strength of the manuscript is that the authors design simulations that closely parallel the experimental 
population design to determine exceptations for change in heterozygsosity and allele frequency 
differentiation. The simulations are very useful in that clearly establish a major role of genetic drift, owing 
in large part to small effective population size, in the differentiation of experimental populations. There are 
a couple of points of clarity needed with regard to the simulations. First, it isn't clear how recombination 
was handled within the simulation, other than an indication that "all recombinations events in the 
simulations were carried out in R with the hyped software package…" (p 11). Presumably some number (0 - 
2?) recombination events per arm generation was included in the simulation, with probability of 
recombination at a location determine by genetic map length. Something like this needs to be stated. Also 
related to the simulations, there are multiple references in the text to empirical results as "deviation from 
simulation" (e.g., p 15) when I suspect what the authors wish to say is that the results differ from neutral 
expectations based on simulations.

The other major weakness of the paper is that relatively little background on the experimental populations 
is provided. There is a long list of citations of prior work on this experimental population, but little 
exposition. A better presentation of the specific target of selection is needed (I infer that selection is on 
improved yield through increased hybrid performance). Ultimately in the Discussion, there is additional 
discussion of the phenotypes, but the presentation is minimal (again pointing to multiple citations) and 
provides little insight into the nature of the selection observed. Better discussion of phenotypes potentially 
targeted by selection is essential if the manuscript is to end, as it does, with a call for further dissection of 
targets of selection in this population.

A third major issue that manuscript seems to be not at all informed by multiple experimental evolution 
studies conducted on Drosophila and other model systems. This contributes to a couple of awkward 
citations of personal communication (p 17 and 19), where ideas about loss of diversity and need for 
replication of experimental populations have been reported in the primary literature. See Drosophila papers 
by Tom Turner and Anthony Long and colleagues.

Finally, the Discussion seems a bit overly long and could potentially be more tightly written.

Minor points: 
- Careful copy editing is needed. Heterozygsoity in the figures and figure captions is referred to variously 
as "H" with and without italics and also as heterozygosity spelled out, sometimes within th esame figure or 
figure caption. Simiarily, Fst needs to be presented in a consistent manner. Also edit citations for 
consistent capitalization and in=text citations for consistent spacing. Also:
On page 7, line 10: A sentence should not begin with a number.
Page 10, line 14: Citation format is incorrect. 
Page 11, line 16: Number of SNPs passing filters should be reported in ‘Materials and Methods’
Figure 2 caption: What is “H combined” and how was it calculated?
Figure 4 and 5: Please include the simulated data (mean + variance) so the pattern is more apparent.

- Figures could be better labeled, for example, simply label panels A & B of Figure 4  as "Genetic Map and 
"Physical Map" and the cycles as "Cycle."

- When discussing the plant material that was used in the analysis (page 7, line 13), two individuals that 
appeared to be mislabeled had their labels switched. However, later in the paragraph, another labeling 
issue was discovered, and these individuals were removed from the analyses. What is the rationale behind 
the differential treatment for these two labeling issues?

- For the integration of the physical and genetic maps, there are two concerns:
1) Do the SNPs on the Maize 50K SNP chip already have reference positions? How were the SNP contextual 
sequences mapped back to the reference genome? If it was a BLAST-type tool, was there an e-value 
threshold that was used to remove low-confidence positions?

2) When integrating the genetic and physical maps, how many SNPs occurred in “mis-mapped blocks?”

- When the cycles of selection shift from 10 individuals to 20 individuals, does this really constitute a 
doubling in effective population size?

- A subheading on heterosis would help better organize the Discussion.

- The Hudson et al. 1992 measure of Fst that the authors report is a sequence-based measure based on 
average pairwise diversity within and between loci. How were the "loci" defined here?

- On page 6, what is the difference in a "found inbred" and a "founder line?" I suspect this distinction will 
be lost on many readers.

- On page 15."consistent with strong genetic drift imposed by selection practices," needs to be reworded.

In summary, this is an interesting manuscript that would benefit from improved clarity of presentation.

\bibliography{references}
\bibliographystyle{geneticsT2}

\end{document}

### 2015-06-30 response.tex

Do the results provide sufficient power to reject the overdominance model? How strong would the signal be likely to be?

\response{blah}

One additional comment: the authors note that heterozygosity increases early in the experiment, and discuss various possible explanations for this. But couldn't directional selection on rare alleles be the driver of this, i.e. early on there will be a rise in allele frequencies, increasing diversity, and then subsequently a loss due to fixations?

\response{blah}

\section*{Reviews}
\subsection*{Reviewer 1}
I really enjoyed reading this manuscript.  I think this is a well done study 
that provides novel insight in to basic processes in breeding of maize.  Many of 
these results have implications for other species as well.  I have a number of 
minor comments below that I think should be addressed but these do not diminish 
my enthusiasm for the manuscript.

@@

In the introduction the authors describe the concept of heterotic groups and 
state that these lines maximize performance and heterosis when crossed.  I would 
caution that this should be clarified to heterosis for yield rather than for all 
traits. Research by other groups (ie Flint-Garcia et al 2009) suggests that 
heterosis can vary for different traits and that the definition of heterotic 
groups is primarily based on yield heterosis, not heterosis for other traits.

\response{blah}

The introduction references several studies that looked at genome-wide 
population structure in breeding programs from the Ross-Ibarra group.  I 
hesitate to even make this suggestion (do to potential flaws) but it might be 
appropriate to also include Jiao et al nat Genetics 2012 here as this study also 
looked at a large dataset of SNPs in modern maize.

\response{While \citet{jiao2012genome} have decided against including citation of the Jiao et al paper, mostly because we are unsure of the veracity of the results (see \citet{lorenz2013phylogenetic}) and concerned by the fact that a wholesale recoding of their genotyping data appeared to have little impact on their conclusions.}

@@

Figure 3 is quite useful and helps to show large physical regions that have lost 
heterozygosity.  The authors discuss the physical-genetic difference.  It would 
be useful to have a supplemental figure that shows these same values (cycle 16 
heterozygosity on y-axis) versus genetic position on x-axis.  I am curious if 
there are ever regions of multiple cM on chromosome arms that become fixed.  
This would be a nice complementary figure.  These are provided for some 
chromosomes in 4B and 5B but would be useful to see genome wide.

\response{blah}

The large regions of fixation are quite intriguing.  Is there any evidence for 
major flowering time or disease resistance QTL in these regions?

\response{blah}

I am somewhat baffled as to why the Fig S1 was put in supplemental materials.  A 
central aspect of this study is this issue of genetic drift relative to 
selection and there are two pages of results text for this section.  However, 
this useful figure is only in supplemental materials.  I would strongly 
encourage moving this figure to the actual manuscript.  If space is an issue 
then only use figu 4 or fig 5 as these are quite redundant with each other in 
terms of conceptual views. 
The legend to fig S1 could use additional detail.  I am guessing that the black 
lines are 95\% CI but this is not clear.  Please explain the black lines.

\response{blah}

@@

Finally, the Discussion seems a bit overly long and could potentially be more tightly written.

\response{blah}

Minor points: 
- Careful copy editing is needed. Heterozygsoity in the figures and figure captions is referred to variously 
as "H" with and without italics and also as heterozygosity spelled out, sometimes within th esame figure or 
figure caption. Simiarily, Fst needs to be presented in a consistent manner. Also edit citations for 
consistent capitalization and in=text citations for consistent spacing. Also:
On page 7, line 10: A sentence should not begin with a number.
Page 10, line 14: Citation format is incorrect. 
Page 11, line 16: Number of SNPs passing filters should be reported in ‘Materials and Methods’
Figure 2 caption: What is “H combined” and how was it calculated?
Figure 4 and 5: Please include the simulated data (mean + variance) so the pattern is more apparent.

\response{blah}

Figures could be better labeled, for example, simply label panels A \& B of Figure 4  as "Genetic Map and 
"Physical Map" and the cycles as "Cycle."

When discussing the plant material that was used in the analysis (page 7, line 13), two individuals that 
appeared to be mislabeled had their labels switched. However, later in the paragraph, another labeling 
issue was discovered, and these individuals were removed from the analyses. What is the rationale behind 
the differential treatment for these two labeling issues?

\response{blah}

For the integration of the physical and genetic maps, there are two concerns:
1) Do the SNPs on the Maize 50K SNP chip already have reference positions? How were the SNP contextual 
sequences mapped back to the reference genome? If it was a BLAST-type tool, was there an e-value 
threshold that was used to remove low-confidence positions?

\response{blah}

@@

2) When integrating the genetic and physical maps, how many SNPs occurred in “mis-mapped blocks?”

\response{blah}

When the cycles of selection shift from 10 individuals to 20 individuals, does this really constitute a 
doubling in effective population size?

\response{blah}

A subheading on heterosis would help better organize the Discussion.

\response{blah}

The Hudson et al. 1992 measure of Fst that the authors report is a sequence-based measure based on 
average pairwise diversity within and between loci. How were the "loci" defined here?

\response{blah}

On page 6, what is the difference in a "found inbred" and a "founder line?" I suspect this distinction will 
be lost on many readers.

\response{blah}

On page 15."consistent with strong genetic drift imposed by selection practices," needs to be reworded.

\response{blah}

### 2015-07-03 response.tex

\maketitle

\section*{Editor's comments}

The reviewers make a number of important points that should be considered in revision. Of particular note, Reviewer 1 highlights the importance of considering the possible effects of purging during the experiment, and this intersects with reviewer 2's comments about lack of details of the experiment and what was being selected, and what the outcome of the experiment was. Could purging, for example, have occurred over the course of the experiment, and did yield improve within lines in addition to increased yield in hybrids? Reviewer 2 also notes a lack of details on the simulation approach; I agree with this, and also thought this was perhaps confusing wording on page 14: "Selections were executed at random".

\response{We thank the editor and reviewers for their time and helpful comments. We have addressed these and other comments specifically in the text below.  Text in the manuscript modified in response to the reviewers has been highlighted blue. }

@@

On the statistical side, candidate selected regions are picked out and effectively 1-sided tests are reported for the rare allele that has increased in frequency. While the authors are reasonably cautious about these cases anyway, at least some attempt to report an experiment-wide p value or FDR for these regions seems useful.

\response{We feel that while an experiment-wide p-value would be important if we were performing a formal genome scan, our purpose here was simply to show that while our overall power to distinguish selection from drift may be low, there are nonetheless individual genomic regions that show patterns of fixation suggestive of selection. We report the \% of simulations for each region that showed lower heterozygosity in simulations, so a simple bonferroni or other calculation could be calculated by a reader concerned we are not sufficiently conservative. \jri{Do we want to do an FDR for the p-values in Figure 3?}}

@@

The abstract could use some editing.  There are a number of references in this 
abstract and the abstract is not well focused on the specific findings of this 
study.

\response{blah}

In the introduction the authors describe the concept of heterotic groups and state that these lines maximize performance and heterosis when crossed.  I would caution that this should be clarified to heterosis for yield rather than for all traits. Research by other groups (ie Flint-Garcia et al 2009) suggests that heterosis can vary for different traits and that the definition of heterotic groups is primarily based on yield heterosis, not heterosis for other traits.

\response{This is a good point and we have added a few words in the intro in the hopes of making this clearer.}

@@

The introduction references several studies that looked at genome-wide 
population structure in breeding programs from the Ross-Ibarra group.  I 
hesitate to even make this suggestion (do to potential flaws) but it might be appropriate to also include Jiao et al nat Genetics 2012 here as this study also looked at a large dataset of SNPs in modern maize.

\response{We had avoided citing \citet{jiao2012genome} because we were  unsure of the veracity of the results (see \citet{lorenz2013phylogenetic}) and concerned by the fact that a wholesale recoding of their genotyping data appeared to have little impact on their conclusions. We have now added a citation, however, as it is certainly true that they performed similar analyses.}

I was quite intrigued by the difference in distance between cycles 4 and then 8-16.  The authors suggest that this could be due to some differences in genetic maintenance of the populations.  I am curious if there were any particular environmental influences during that time period that may have led to stronger selection or if there were any differences in agronomic practices that have occurred during the long-term experiment (ie irrigation, fertilization, plant density, etc).  it would be helpful to know if any of these factors may have contributed.

\response{\jri{Jode? Justin? I don't have the answer here, but suspect it is ``no''}}

@@

It was a bit surprising to me that there was no differentiation among the founder inbreds of the two populations.  If PCA was run only on the founders (without any other lines) do you get separation of the two groups or not?

\response{\jri{Justin, have you done this? Easy to do?}}

Figure 3 is quite useful and helps to show large physical regions that have lost heterozygosity.  The authors discuss the physical-genetic difference.  It would be useful to have a supplemental figure that shows these same values (cycle 16 heterozygosity on y-axis) versus genetic position on x-axis.  I am curious if there are ever regions of multiple cM on chromosome arms that become fixed.  This would be a nice complementary figure.  These are provided for some chromosomes in 4B and 5B but would be useful to see genome wide.

\response{Heterozygosity plotted on a genetic map was shown in the original supplemental figure 2 (which the reviewer pointed out we failed to reference).  We have switched this figure for the original figure 2. As shown in the new figure 3, there are a few regions on chromosome arms that have been fixed and span multiple cM. \jri{do we need to note this in the main text? is it an intesting observation, maybe implying rapid selection?}}

@@

The large regions of fixation are quite intriguing.  Is there any evidence for major flowering time or disease resistance QTL in these regions?

\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? Or say ``geez this is hard because QTL are big and we're not sure we'd believe the results?''}

@@

I am somewhat baffled as to why the Fig S1 was put in supplemental materials.  A central aspect of this study is this issue of genetic drift relative to selection and there are two pages of results text for this section.  However, this useful figure is only in supplemental materials.  I would strongly encourage moving this figure to the actual manuscript.  If space is an issue then only use figu 4 or fig 5 as these are quite redundant with each other in terms of conceptual views. 
The legend to fig S1 could use additional detail.  I am guessing that the black lines are 95\% CI but this is not clear.  Please explain the black lines.

\response{Following the reviewer's suggestion we have moved this figure to the main text; the figure legend now explains the black lines. The old figures 4 and 5 have been combined into a single multi-panel plot.}

@@

Figure S2 is not ever referenced in the text as far as I could find.  The data 
table of the same results is referenced but I don’t think the figure is.

\response{Figure S2 and Figure 3 have been switched as described above; the supplmental figure is now referenced in the caption to the (new) Figure 3.}

### 2015-07-03 discussion.tex
working on response

Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah19861985, 1986; troyer1999background}. 
B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
We examined these three inbreds at the pericentromeric regions listed in Table \ref{tabfix} \jri{correct?}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.  
Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.

>>>>

Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
We examined these three inbreds at the pericentromeric regions listed in Table \ref{tabfix} \jri{correct?}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.  
Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.

### 2015-07-03 intro.tex
working on response

\section*{Introduction}
Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
Hybrid maize, first developed in the early 20th century by \jri{East and Shull?}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}. 
The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.

Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}. 
Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}. 
These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical}. 
Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}. 
The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.

Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
\jri{add some text explaining recurrent selection}
This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah19861985}, the formation of modern heterotic groups \citep{troyer1999background; senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
\jri{something about heterosis}

>>>>

\section*{Introduction}
Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
Hybrid maize, first developed in the early 20th century \citep{crow1998}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}. 
The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) \rev{for yield} when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.

Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}. 
Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}. 
These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}. 
Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}. 
The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.

Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
\jri{add some text explaining recurrent selection}
This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
\jri{something about heterosis}
